---
title: "Exploration to the influences and relation of different variables to the IMBD score of moviest"
author: "Sicong He & Ruiyang Dai"
date: "2020/8/2"
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default
urlcolor: cyan
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction
Both members in our team has a strong interest in movies, thus we decided to look for a dataset about movies as our final project. After a extensive search, we choose this dataset with IMDB scores of 5000 movies. In this project, we are going to explore the potential relationships between all other variables like budget, year of publication, director, etc. and the IMDB score of the movie. Using IMDB scores as the response, we expect to build a good model that could help us predict the potential imdb scores of the new releasing movies.  
  
The information of the data, and the variables we want to use are listed below:  
The dataset collected data of 28 variables from 5000 movies.   
The dataset contains variables such as director name, budget, imdb scores, year of publication, and etc.

The variables we are about to use in this project: 

`Color`: ` Black and White` and `Color`, depends on the color of the movie

`num_critic_for_reviews`: number of critics that have written review articles about the movie

`duration`: total length of the movie

`director_facebook_likes`: likes the director get from facebook about the movie on facebook

`actor_1_facebook_likes`: number of likes the actor who play the leading role in the movie on facebook

`gross`: the money earned from the movie, in dollars, globally.

`num_voted_users`: number of users who vote for the movie on imdb website
‘cast_total_facebook_likes: Total facebook likes for this movie.

`num_user_for_reviews`: Number of users for reviews.

`Language`: English and other.

`content_rating`: Level of content so parents may decide whether the movie is suitable for their children, PG, PG-13, R, and others.

`budget`: budget

`title_year`: year of publication

`imdb_score`: scores that given by IMDB


Source: https://www.kaggle.com/carolzhangdc/imdb-5000-movie-dataset

# Methods
```{r message=FALSE, warning=FALSE}
library(readr)
library(forcats)
library(faraway)
library(lmtest)
```

```{r}
movie = read.csv("movie_metadata.csv")
movie = movie[,c("color", "num_critic_for_reviews", "duration", "director_facebook_likes", "actor_1_facebook_likes", "gross", "num_voted_users", "num_user_for_reviews", "language", "content_rating", "budget", "title_year", "imdb_score")]
movie$language = fct_other(movie$language, keep = "English", other_level = "Other") #simplify movie languages into only 2 factors
movie$content_rating = fct_other(movie$content_rating, keep = c("R", "PG-13", "PG"), other_level = "Other") #simplify content ratings into only 4 factors. 
movie = movie[complete.cases(movie), ] #delete observations with NA
train = sample(1:3886, 800)
is.factor(movie$color)
movie$color = as.factor(movie$color)
is.factor(movie$language)
movie$language = as.factor(movie$language)
is.factor(movie$content_rating)
movie$content_rating = as.factor(movie$content_rating)
is.factor(movie$content_rating)
train_data = movie[train,]
n = 800
test_data = movie[-train,]
```
```{r, fig.height=20, fig.width=20, eval=FALSE}
plot(movie, col = "dodgerblue")
```
Therer seems to be relationships existing between IMDB scores with all the variables, and we observed a stong colinearity between duration and number of reviews, and there seems to exist some relationship between gross~all other numeric factors
## Model selection by AIC and BIC

Firstly, we decided to use a BIC model to begin our variable selection, using stepwise search checks going both backwards and forwards at every step. Then, we set the full interactive model as an end point. 

```{r}
full_add_model = lm(imdb_score ~ ., data = train_data)
BIC_model  = step(full_add_model, scope = imdb_score ~ .^2, direction = "both", k = log(n), trace = 0)
```

Here are result of the BIC model:
```{r}
coef(BIC_model)

```

## Model Diagnostics -- Check constant variance assumption

We create functions that could help us check constant variance assumption
```{r}
# A test for constant variance
const_var_test = function(model) {
  p = bptest(model)$p.value[[1]]
  print(c("p-value is", signif(p, 3)))
  ifelse(p > 0.05, "Do not reject H0, The errors have constant variance about the true model."
         , "Reject H0, The errors have non-constant variance about the true model.")
}
# A function that can help to draw a residual plot
resid_plot = function(model) {
  plot(fitted(model), resid(model))
  abline(h = 0)
}
```




Apply test functions to  BIC model

```{r}
const_var_test(BIC_model)
resid_plot(BIC_model)
```

## Model Diagnostics -- Check normality assumption

We create functions that could help us check normality assumption
```{r}
# A test for Normality
normality_test = function(model) {
  p = shapiro.test(resid(model))$p.value
  print(c("p-value is", signif(p, 3)))
  ifelse(p > 0.05, "Do not reject H0, We assume the data were sampled from a normal distribution."
         , "Reject H0, We do not assume the data were sampled from a normal distribution.")
}
# A function that can help to draw a normal Q-Q plot
qq_plot = function(model) {
  qqnorm(resid(model))
  qqline(resid(model), col = "dodgerblue", lwd = 2)
}
```

Apply these functions on our  BIC model

```{r}
normality_test(BIC_model)
qq_plot(BIC_model)
shapiro.test(resid(BIC_model))
```
We observed an obvious pattern in Q-Q plot that the data does not follows a normal distribution and a fat tailat the left. Thus we decided to try a data transformation.
```{r}
model_BIC = lm(imdb_score~color + num_critic_for_reviews + duration + gross + num_voted_users + language + content_rating + budget 
   + title_year + duration:num_voted_users + duration:content_rating + num_critic_for_reviews:content_rating + language:budget, data = train_data)
```
we first tried a log transformation:  

```{r}
log_imdb = lm(imdb_score ~ color + num_critic_for_reviews + duration + gross + num_voted_users + language + content_rating + budget 
   + title_year+ duration:num_voted_users + duration:content_rating + num_critic_for_reviews:content_rating + language:budget + log(num_critic_for_reviews) + log(duration) + log(num_user_for_reviews) + log(budget) + log(title_year) + log(gross) + log(num_voted_users) , data = train_data)
back_BIC = step(log_imdb, trace = 0, k = log(n))
coef(back_BIC)
```

and selected the log variables BIC test remains to perform a model:  
```{r}
log_movie2 = lm(imdb_score ~ color +duration +gross +  log(num_critic_for_reviews) +  log(num_user_for_reviews) +content_rating+
                  language + log(budget) + log(num_voted_users) +log(num_voted_users) +  duration:num_voted_users + duration:content_rating
                +num_critic_for_reviews:content_rating + language:budget, data = train_data)
```
```{r}
normality_test(log_movie2)
const_var_test(log_movie2)
```


```{r}
qq_plot(log_movie2)
```
we see no obvious improvement in the normailty and equal variance assumptions.  

Then, we decided to try $1/x$:  
we first perform it on the response:
```{r}
fraction_movie = lm(1/(imdb_score)~color + num_critic_for_reviews + duration + gross + num_voted_users + language + content_rating + budget 
   + title_year + duration:num_voted_users + duration:content_rating + num_critic_for_reviews:content_rating + language:budget, data = train_data)
```

```{r}
normality_test(fraction_movie)
const_var_test(fraction_movie)
summary(fraction_movie)$adj.r.squared
```
```{r}
qq_plot(fraction_movie)
resid_plot(fraction_movie)
```

we observed that using a fraction method does improve the assumption of model significantly. With the qqplot plot we observed an obvious "fat tail" on the right.  
To deal with the tail, we first analyze our trained dataset to remove the unusual observations:   
first we check for observations that are influrential:  
```{r}
outliers=subset(train_data, cooks.distance(fraction_movie) > 4/length(cooks.distance(fraction_movie)))
newdata = subset(train_data, cooks.distance(fraction_movie) <= 4/length(cooks.distance(fraction_movie)))
outliers
```
after removing the influential outliers, we fit the new data into our model again:  
```{r}
fraction_movie_new = lm(1/(imdb_score)~color + num_critic_for_reviews + duration + gross + num_voted_users + language + content_rating + budget 
   + title_year + duration:num_voted_users + duration:content_rating + num_critic_for_reviews:content_rating + language:budget, data = newdata)
```
and perform test to check its assumptions for normality and equal varia:  
```{r}
normality_test(fraction_movie_new)
const_var_test(fraction_movie_new)
```
```{r, fig.height=8, fig.width=10}
par(mfrow = c(1,2))
qq_plot(fraction_movie_new)
resid_plot(fraction_movie_new)
```
We can still observed the cluster in our residual plot, thus we decided to try to transform our data, to improve the performance.  

We first tried to square the response, but does not see significant improvement
```{r}
fraction_movie_new1 = lm(I(1/(imdb_score)^2)~color + num_critic_for_reviews + duration + gross + num_voted_users + language + content_rating + budget 
   + title_year + duration:num_voted_users + duration:content_rating + num_critic_for_reviews:content_rating + language:budget, data = newdata)
```
```{r}
normality_test(fraction_movie_new1)
const_var_test(fraction_movie)
```
```{r}
par(mfrow = c(1,2))
qq_plot(fraction_movie)
resid_plot(fraction_movie)
```
no obvious improvement for the model.  
Then we try log transformation to our response:  
```{r}
fraction_movie_log = lm(log(1/(imdb_score))~color + num_critic_for_reviews + duration + gross + num_voted_users + language + content_rating + budget 
   + title_year + duration:num_voted_users + duration:content_rating + num_critic_for_reviews:content_rating + language:budget, data = newdata)
```
```{r}
normality_test(fraction_movie_log)
const_var_test(fraction_movie_log)
```
```{r, fig.height=8, fig.width=10}
par(mfrow = c(1,2))
qq_plot(fraction_movie_log)
resid_plot(fraction_movie_log)
```
we do not see improve in performance for any of the assumptions.  
then, we try the backward and stepwise BIC method for our fraction model to see if we can eliminate some factors:  
```{r}
new_n  = 771
frac_back_BIC = step(fraction_movie_new, trace = 0, k = log(new_n))
frac_step_BIC = step(fraction_movie_new, scope = imdb_score ~ .^2, direction = "both", k=log(new_n), trace = 0)
```
```{r}
length(coef(frac_back_BIC)) > length(coef(frac_step_BIC))
```

```{r}
summary(frac_back_BIC)$adj.r.squared
summary(frac_step_BIC)$adj.r.squared
```
We choose the model with backward BIC.

```{r}
normality_test(frac_back_BIC)
const_var_test(frac_back_BIC)
```
```{r, fig.height=8, fig.width=10}
par(mfrow = c(1,2))
qq_plot(frac_back_BIC)
resid_plot(frac_back_BIC)
```

```{r}
plot(frac_back_BIC)
```

from the plot above we can see that the data marked 
```{r}
rm = which(rownames(newdata) == "1565")
names(frac_back_BIC)
plot(newdata[-515,])
```



